---
title: "Simple Linear Regression"
subtitle: "Inference"
author: "Dr. Maria Tackett"
output:
  xaringan::moon_reader:
    mathjax: "https://cdn.bootcss.com/mathjax/2.7.1/MathJax.js?config=TeX-MML-AM_HTMLorMML"
    css: "slides.css"
    logo: "img/introregression-sticker.png"
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: 4:3
      slideNumberFormat: "%current%"
    seal: false #make custom title slide
editor_options: 
  chunk_output_type: console
---

```{r child = "setup.Rmd"}
```

class: title-slide 

<br><br>

# Simple Linear Regression 
## Inference

<br><br><br>

### Dr. Maria Tackett

---

## Topics 

--


  
--



```{r packages, echo = F}
library(tidyverse)
library(broom)
library(knitr)
library(kableExtra)
library(patchwork)
library(fivethirtyeight)
```

```{r data}
movie_scores <- fandango %>%
  rename(critics = rottentomatoes, 
         audience = rottentomatoes_user)
```

---

## Movie ratings data 

The data set contains the "Tomatometer" score (.term[critics]) and audience score (.term[audience]) for 146 movies rated on rottentomatoes.com. 

```{r}
ggplot(data = movie_scores, mapping = aes(x = critics, y = audience)) +
  geom_point(alpha = 0.5) + 
  labs(x = "Critics Score" , 
       y = "Audience Score") +
  theme_bw() +
  theme(axis.title=element_text(size=14))
```


---

## The model

$$\color{red}{\hat{\text{audience}} = 32.316 + 0.519 \times \text{critics}}$$

```{r}
model <- lm(audience ~ critics, data = movie_scores)
```

```{r}
model %>%
  tidy() %>%
  kable(format = "html", digits = 3)
```


```{r}
ggplot(data = movie_scores, mapping = aes(x = critics, y = audience)) +
  geom_point(alpha = 0.5) + 
  geom_smooth(method = "lm", color = "red", se = FALSE) +
  labs(x = "Critics Score" , 
       y = "Audience Score") +
  theme_bw() +
  theme(axis.title=element_text(size=14))
```

---

## Questions of interest

--

Is there truly linear relationship between `critics` and `audience`, or is the relationshp we observed $(\hat{\beta}_1 = 0.519)$ due to random chance? 

<br> 

--

What is a plausible range of values of $\beta$, the true population slope for `critics`? 

---

class: middle

### Is there truly linear relationship between `critics` and `audience`, or is the relationshp we observed $(\hat{\beta}_1 = 0.519)$ due to random chance? 


---

## Outline of a hypothesis test

`r emo::ji('one')` State the hypotheses.

`r emo::ji('two')` Calculate the test statistic. 

`r emo::ji('three')` Calculate the p-value.

`r emo::ji('four')` State the conclusion. 

---

## `r emo::ji('one')` State the hypotheses

```{r}
model %>%
  tidy() %>%
  kable(format = "html", digits = 3)
```

<br>

.pull-left[
.small-box[
$$\large{\begin{aligned}& H_0: \beta_1 = 0\\& H_a: \beta_1 \neq 0\end{aligned}}$$
]
]

---

## `r emo::ji('one')` State the hypotheses

```{r}
model %>%
  tidy() %>%
  kable(format = "html", digits = 3)
```

<br>

.pull-left[
.small-box[
$$\large{\begin{aligned}& H_0: \beta_1 = 0\\& H_a: \beta_1 \neq 0\end{aligned}}$$
]
]

.pull-right[
<font color = "white">place-holder</font>

.big[.vocab[Null hypothesis]]


]

---

## `r emo::ji('one')` State the hypotheses

```{r}
model %>%
  tidy() %>%
  kable(format = "html", digits = 3)
```

<br>

.pull-left[
.small-box[
$$\large{\begin{aligned}& H_0: \beta_1 = 0\\& H_a: \beta_1 \neq 0\end{aligned}}$$
]
]

.pull-right[
<font color = "white">place-holder</font>
 
.big[.vocab[Null hypothesis]]

.big[.vocab[Alternative hypothesis]]

]

---

## `r emo::ji('two')` Calculate the test statistic

```{r}
model %>%
  tidy() %>%
  kable(format = "html", digits = 3)
```

<br> 

.eq[
$$\text{test statistic} = \frac{\text{Estimate} - \text{Hypothesized}}{\text{Standard error}}$$
]

---

## `r emo::ji('two')` Calculate the test statistic

```{r}
model %>%
  tidy() %>%
  kable(format = "html", digits = 3) %>%
  row_spec(2, background = "#dce5b2")
```

<br> 

--

.pull-left[
.eq[
$$t = \frac{{\hat{\beta}_1} - 0}{\text{SE}_{\hat{\beta}_1}}$$
]
]

--

.pull-right[

.small-box-work[
$$\begin{aligned}t &= \frac{0.5187 - 0}{0.0345}\\
&= \mathbf{15.03}\end{aligned}$$
]
]

---

### 3. Calculate the p-value

<font class="vocab">p-value</font> is calculated from a $t$ distribution with $n-2$ degrees of freedom

.alert[
$$\text{p-value} = P(t \geq |\text{test statistic}|)$$
]

```{r echo = F}
ggplot(data.frame(x = c(-4, 4)), aes(x)) +
  stat_function(fun = dt, args =list(df =144)) +
  stat_function(fun = dt,   args =list(df =144),
                xlim = c(-4, -1.5),
                geom = "area", fill = "navy") + 
    stat_function(fun = dt,   args =list(df =144),
                xlim = c(1.5, 4),
                geom = "area", fill = "navy")
```

---


class: middle, center

.instructions[
**Write the general definition of the p-value for tests of $\beta_1$.**
]

---

### 4. State the conclusion 

|  Magnitude of p-value |             Interpretation            |
|:---------------------:|:-------------------------------------:|
| p-value < 0.01        | strong evidence against $H_0$         |
| 0.01 < p-value < 0.05 | moderate evidence against $H_0$       |
| 0.05 < p-value < 0.1  | weak evidence against $H_0$           |
| p-value > 0.1         | effectively no evidence against $H_0$ |
<br> 
<br>

**Notes:** 

- These are general guidelines. The strength of evidence depends on the context of the problem.
- Don't just rely on the reject/fail to reject conclusion from the hypothesis test to draw conclusions about the relationship between $x$ and $y$. Use the EDA and confidence intervals to provide more context about the implications of your results.

---

### Movie data: Hypothesis test for $\beta_1$ 

```{r}
model %>%
  tidy() %>%
  kable(format = "markdown", digits = 3)
```

.instructions[
a. State the hypotheses in (1) words and (2) statistical notation.  

b. What is the meaning of the test statistic in the context of the problem? 

c. What is the meaning of the p-value in the context of the problem?

d. State the conclusion in context of the problem.
]
---

class: middle, center

### What is a plausible range of values of the true population slope for `critics`?

---

### General form of the confidence interval

- Let <font class="vocab">SE</font> be the standard error of the statistic used to estimate the parameter of interest, then the general form of the confidence interval is

.alert[
$$\text{ Estimate} \pm \text{ (critical value) } \times \text{SE}$$
]
- *Note*: The critical value is determined by the distribution of the estimate (statistic) and the confidence level

- For the regression slope: 
    - $\hat{\beta}_1$ is the statistic used to estimate the parameter, $\beta_1$ 
    - We will write the confidence interval as 
    $$\mathbf{\hat{\beta}_1 \pm t^* SE(\hat{\beta}_1)}$$
    
---

### Confidence interval for $\beta_1$

- The confidence interval for the regression slope is 

.alert[
$$\mathbf{\hat{\beta}_1 \pm t^* SE(\hat{\beta}_1)}$$
]

- $t^*$ is the critical value associated with the confidence level.
  + It is calculated from a $t$ distribution with $n-2$ degrees of freedom
  
- $SE(\hat{\beta}_1)$ is the standard error for the slope 

$$SE(\hat{\beta}_1) = \sqrt{\frac{\hat{\sigma}^2}{\sum\limits_{i=1}^n (x_i - \bar{x})^2}} \hspace{2.5mm} = \hspace{2.5mm} \hat{\sigma}\sqrt{\frac{1}{(n-1)s_X^2}}$$

---

### What is $\hat{\sigma}$?

- Recall, the residual is the difference between the observed response the predicted response (the estimated mean) 
    - The residual for the ith observation, $(x_i, y_i)$, is
    
    $$e_i = y_i - (\hat{\beta}_0 + \hat{\beta}_1 x_i)$$ 

- The <font class = "vocab">Residual Standard Error</font> is the estimate of variation about the regression line
    - Also known as the **Root Mean Square Error (RMSE)**

.alert[
$$\hat{\sigma} = \sqrt{\frac{1}{n-2}\sum\limits_{i=1}^{n} e_i^2}$$
]
---

### Why *t*? 

.alert[
$$\hat{\beta}_1 \sim N \Bigg(\beta_1, \sigma\sqrt{\frac{1}{(n-1)s_X^2}} \Bigg)$$
]

- We don't know $\sigma$, so we use its estimate $\hat{\sigma}$ in our calculations. Therefore, we use the *t* distribution when we calculate the confidence interval (and conduct hypothesis tests) to account for the extra variability that's been introduced (same reason we use *t* when doing inference for a mean)

- The critical value $t^*$ is calculated from the *t(n-2)* distribution - the *t* distribution with *n-2* degrees of freedom.

---

### Movies data: Critical value

```{r}
qt(0.975, 144)
```

```{r echo = F}
ggplot(data.frame(x = c(-4, 4)), aes(x)) +
  stat_function(fun = dt, args =list(df =144)) +
  stat_function(fun = dt,   args =list(df =144),
                xlim = c(-1.976575, 1.976575),
                geom = "area", fill = "navy") + 
  geom_text(x=0, y=0.15, label="0.95", color = "white", size = 10) +
  geom_text(x = -2.2, y = 0.07, label = "-1.977", size = 3) + 
  geom_text(x =  2.2, y = 0.07, label = "1.977", size = 3)
```

---
### Calculating the 95% CI for $\beta_1$

```{r echo = F}
n <- nrow(movie_scores)
var.x <- var(movie_scores$critics)
sigma <- glance(model)$sigma
beta1 <- tidy(model)$estimate[2]
crit.val <- qt(0.975, n-2)
summaries <- tibble(n = n, 
                    var.x = var.x,
                    sigma = sigma,
                    beta1 = beta1, 
                    crit.val = crit.val)
summaries %>% kable(format = "markdown", digits = 3)
```

.instructions[
Write the equation for the 95% confidence interval for $\beta_1$, the coefficient (slope) of `critics`. 
]

---

### Interpretation 

```{r}
model %>%
  tidy(conf.int=TRUE) %>%
  kable(format = "markdown", digits = 3)
```

.instructions[
Interpret the 95% confidence interval for $\beta_1$, the coefficient (slope) of critics.
]





---

## Recap
--


--


